# Title: "Reproducible Research_Assignment 1"  
  Author: Harsha Chandapuram


# reading the data
```{r}
setwd("C:/Users/Harsha/Desktop/RepData_PeerAssessment1")
md <- read.csv("activity.csv",header = T)
head(md)
```

# changing the date variable into date
```{r}
md$date <- as.Date(md$date)

```

# Sorting data by removing NAs
```{r}
md1 <- md[!is.na(md$steps),]
head(md1)
```

#summarizing total number of steps taken each day
```{r}
library("plyr",lib.loc = "C:/Users/Harsha/Desktop/rpackages")
tot <- ddply(md1,.(date),summarise,steps = sum(steps,na.rm = T))
head(tot)

```

# histogram of total number of steps taken each day (after removing NAs)

```{r}
library("ggplot2",lib.loc = "C:/Users/Harsha/Desktop/rpackages")
library("labeling",lib.loc = "C:/Users/Harsha/Desktop/rpackages")
a <- ggplot(tot,aes(tot$steps))+geom_histogram(bins = 20,width=0.5,fill="white",col="red")
b <- a + theme(axis.text.x=element_text(angle = 90,hjust = 1,vjust=0.5))
print(b)
```

# mean and median of total number of steps taken per day (after removing NAs.)
```{r}
mean <- mean(tot$steps)
mean
median <- median(tot$steps)
median
```

# Time-series plot of 5min-interval Vs avg.steps per inteval across all days (after removing NAs)
```{r}
fivemin <- ddply(md1,.(interval),summarise,steps = mean(steps,na.rm = T))
base <- ggplot(fivemin,aes(fivemin$interval,y=fivemin$steps))
base <- base + geom_line(stat="identity")+scale_x_continuous(breaks=seq(0,2355,50))
print(base)
```

# Max. no. of steps in an interval across all days 
```{r}
m5m <- fivemin[which.max(fivemin$steps),]
head(m5m)
```

# missing value rows in the data set
```{r}
setwd("C:/Users/Harsha/Desktop/RepData_PeerAssessment1")
mmd <- read.csv("activity.csv",header = T)
missing <- length(mmd$steps[is.na(mmd$steps)])
missing

```

# filling in the missing values.
```{r}
for(i in 1:nrow(fivemin)){
if(any(is.na(md$steps[md$interval==fivemin$interval[i]]))){
md$steps[(md$interval==fivemin$interval[i])][is.na(md$steps[(md$interval==fivemin$interval[i])])] <- fivemin$steps[i]
                }
        }

```

# new totals after missing data filled in
```{r}
ntot <- ddply(md,.(date),summarise,steps=sum(steps,na.rm = T))
head(ntot)

```

# histogram of total number of steps taken each day (After filled in Missing data)
```{r}
newa <- ggplot(ntot,aes(ntot$steps))+geom_histogram(bins = 20,width=0.5,fill="white",col="red")
newa <- newa + theme(axis.text.x=element_text(angle = 90,hjust = 1,vjust=0.5))
print(newa)
```

# mean and median of total number of steps taken per day (after filled in Missing data)
```{r}
mean2 <- mean(ntot$steps)
mean2
median2 <- median(ntot$steps)
median2

```

# changing the dataset into weekdays and weekends
```{r}
library("chron",lib.loc = "C:/Users/Harsha/Desktop/rpackages")
if(any(is.weekend(md$date))){
md["days"] <- ifelse(is.weekend(md$date),"Weekend","Weekday")
}
newmd <- md[c(1,3,4)]
newmd<- ddply(newmd,.(days,interval),summarise,steps=mean(steps))
head(newmd)

```

# plotting panel plot of weekdays and weekends
```{r}
a <- ggplot(newmd, aes(interval,steps))+geom_line(color="blue")+facet_wrap(~days, ncol=1)
print(a)

```





